---
title: "INTRuST CM — H1a Supplementary: CTQ and White Matter Microstructure"
author: "Luisa Berger"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

# Document Overview

This document contains supplementary regression analyses for
**H1a: CTQ total score and white matter (WM) microstructure**.

| Table | Analysis | Purpose |
|-------|----------|---------|
| S2  | Additional covariates (race; race + alcohol) | Robustness to demographic confounds |
| S5  | HC3 SEs + BCa bootstrap 95% CIs | Robustness to heteroscedasticity |
| S6  | Excluding PTSD-only subgroup | Sensitivity to influential subgroup |
| S12 | CTQ subscale analyses (MD + AD) | Exploring which subtype drives main effects |

**Reproduce this document:** open `intrust_cm.Rproj` and click *Knit*, or run
`rmarkdown::render(here::here("code/h1a_associations_ctq_white_matter_supp.Rmd"))`.

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

suppressPackageStartupMessages({
  library(here)
  library(tidyverse)
  library(broom)
  # HC3 robust SEs and BCa bootstrap
  library(sandwich)
  library(lmtest)
  library(boot)
  # Table rendering (all tables use gt for HTML output)
  library(gt)
})

set.seed(1234)

# ---- Data --------------------------------------------------------------------

dat <- readr::read_csv(
  here("data", "analysis_simulated.csv"),
  show_col_types = FALSE,
  name_repair    = "minimal"
)

dat$ctq_totalscore <- as.numeric(dat$ctq_totalscore)
dat$age             <- as.numeric(dat$age)
dat$gender          <- as.numeric(dat$gender)
dat$race            <- as.numeric(dat$race)
dat$audit_10_score  <- as.numeric(dat$audit_10_score)

# ---- Output directory --------------------------------------------------------
# HTML tables are saved under output/tables/supplementary/.
dir.create(here("output", "tables", "supplementary"),
           showWarnings = FALSE, recursive = TRUE)
```

```{r define_lookup, include=FALSE}
# ==============================================================================
# Tract ordering and column-name conventions.
# ==============================================================================

# Order of the 23 pre-selected WM tracts
tract_stems <- c(
  "t_af_left",      "t_af_right",
  "t_cb_left",      "t_cb_right",
  paste0("t_cc", 1:7),
  "t_ilf_left",     "t_ilf_right",
  "t_ioff_left",    "t_ioff_right",
  "t_slf.i_left",   "t_slf.i_right",
  "t_slf.ii_left",  "t_slf.ii_right",
  "t_slf.iii_left", "t_slf.iii_right",
  "t_uf_left",      "t_uf_right"
)

# Column-name suffixes for each diffusion metric
metric_suffixes <- list(
  FA = ".fa1.mean",
  MD = ".tensor1.meandiffusivity.mean",
  AD = ".tensor1.maxeigenvalue.mean",
  RD = ".tensor1.mineigenvalue.mean"
)

# Tract labels
tract_pretty <- c(
  "AF left",      "AF right",
  "CB left",      "CB right",
  paste0("CC ", 1:7),
  "ILF left",     "ILF right",
  "IoFF left",    "IoFF right",
  "SLF I left",   "SLF I right",
  "SLF II left",  "SLF II right",
  "SLF III left", "SLF III right",
  "UF left",      "UF right"
)
names(tract_pretty) <- paste0("Y", seq_along(tract_pretty))

# All expected WM columns must be present in dat.
all_wm_cols <- as.vector(outer(tract_stems, unlist(metric_suffixes), paste0))
missing_wm  <- setdiff(all_wm_cols, names(dat))
if (length(missing_wm) > 0) {
  stop("define_lookup: the following expected WM columns are absent from dat:\n  ",
       paste(missing_wm, collapse = "\n  "))
}

# Check predictor, grouping, and additional covariate columns.
stopifnot(
  "Required columns missing from dat — check analysis_simulated.csv" =
    all(c("ctq_totalscore", "age", "gender",
          "race", "audit_10_score",
          "hc.ptsd.mtbi.groups.4groups",
          "ctq_physneg_score") %in% names(dat))
)
```

```{r helpers, include=FALSE}
# ==============================================================================
# Helper functions shared across all analysis sections.
# ==============================================================================

# strip0() ─────────────────────────────────────────────────────────────────────
# Remove the leading zero from a formatted number, including negatives.
# e.g. "0.123" -> ".123",  "-0.123" -> "-.123"
strip0 <- function(x) gsub("(?<=^|-)0\\.", ".", x, perl = TRUE)


# fit_ols_wm(): multivariate OLS + standardised model for WM tract outcomes.
# Returns list(out = CTQ-row tidy table, beta = standardised betas).
fit_ols_wm <- function(values, data, predictor, p_fdr_col) {

  rhs    <- paste(c(predictor, "age", "gender"), collapse = " + ")
  rhs_sc <- paste(
    paste0("scale(", c(predictor, "age", "gender"), ")"),
    collapse = " + "
  )

  lm_fit          <- lm(as.formula(paste("values ~", rhs)), data = data)
  tt              <- broom::tidy(lm_fit)
  tt[[p_fdr_col]] <- p.adjust(tt$p.value, method = "fdr")

  out <- tt |>
    dplyr::filter(term == predictor) |>
    dplyr::mutate(
      p.value      = formatC(p.value,               digits = 3, format = "f"),
      !!p_fdr_col := formatC(.data[[p_fdr_col]], digits = 3, format = "f")
    ) |>
    dplyr::select(response, statistic, p.value, dplyr::all_of(p_fdr_col))

  lm_sc <- lm(as.formula(paste("scale(values) ~", rhs_sc)), data = data)
  beta  <- broom::tidy(lm_sc) |>
    dplyr::filter(term == paste0("scale(", predictor, ")")) |>
    dplyr::mutate(estimate = formatC(estimate, digits = 3, format = "f")) |>
    dplyr::select(response, estimate)

  invisible(list(out = out, beta = beta))
}


# fit_hc3_bca(): per-tract HC3 SEs and BCa bootstrap CIs on standardised betas.
# Returns tibble with p_HC3, pFDR_HC3, HC3_lo/hi_std, BCa_lo/hi_std.
fit_hc3_bca <- function(values, ynames, data, predictor, R_boot = 5000) {

  rhs <- paste(c(predictor, "age", "gender"), collapse = " + ")

  # (A) HC3 p-values: unstandardised model, BH FDR across all terms × tracts --
  hc3_all <- purrr::map_dfr(seq_along(ynames), function(i) {
    y   <- as.numeric(values[, i])
    fit <- lm(as.formula(paste("y ~", rhs)), data = data)
    Vh  <- sandwich::vcovHC(fit, type = "HC3")
    ct  <- broom::tidy(
      lmtest::coeftest(fit, vcov. = Vh, df = stats::df.residual(fit))
    )
    dplyr::mutate(ct, response = ynames[i], .before = 1) |>
      dplyr::select(response, term, p.value) |>
      dplyr::rename(p_HC3 = p.value)
  })
  hc3_all$pFDR_HC3_all <- p.adjust(hc3_all$p_HC3, method = "BH")

  hc3_p <- hc3_all |>
    dplyr::filter(term == predictor) |>
    dplyr::select(response, p_HC3, pFDR_HC3 = pFDR_HC3_all)

  # (B) Standardised HC3 CIs + BCa bootstrap for predictor only ----------------
  hc3_ci <- purrr::map_dfr(seq_along(ynames), function(i) {
    y  <- as.numeric(values[, i])
    df <- data.frame(
      y_s      = scale(y)[, 1],
      ctq_s    = scale(data[[predictor]])[, 1],
      age_s    = scale(data$age)[, 1],
      gender_s = scale(data$gender)[, 1]
    )
    fit <- lm(y_s ~ ctq_s + age_s + gender_s, data = df)
    Vh  <- sandwich::vcovHC(fit, type = "HC3")
    ci  <- lmtest::coefci(fit, vcov. = Vh, df = stats::df.residual(fit))["ctq_s", ]

    stat_fun <- function(d, idx) {
      unname(coef(
        lm(y_s ~ ctq_s + age_s + gender_s, data = d[idx, , drop = FALSE])
      )["ctq_s"])
    }
    bobj <- boot::boot(df, statistic = stat_fun, R = R_boot)
    bca  <- tryCatch(
      boot::boot.ci(bobj, type = "bca", index = 1)$bca[4:5],
      error = function(e) c(NA_real_, NA_real_)
    )

    tibble::tibble(
      response   = ynames[i],
      HC3_lo_std = ci[1], HC3_hi_std = ci[2],
      BCa_lo_std = bca[1], BCa_hi_std = bca[2]
    )
  })

  invisible(dplyr::left_join(hc3_p, hc3_ci, by = "response"))
}


# build_hc3_table() ────────────────────────────────────────────────────────────
#
# `ols_out`      : output of fit_ols_wm()$out for this metric.
# `ols_beta`     : output of fit_ols_wm()$beta for this metric.
# `hc3`          : output of fit_hc3_bca() for this metric.
# `tract_labels` : named vector (Y1..Y23 -> readable names) for row labels
#                  and ordering.
# `p_fdr_col`    : name of the FDR column in ols_out, e.g. "p.fdr.fa".
build_hc3_table <- function(ols_out, ols_beta, hc3, tract_labels, p_fdr_col) {
  ols_out |>
    dplyr::left_join(ols_beta, by = "response") |>
    dplyr::left_join(hc3,      by = "response") |>
    dplyr::mutate(
      Tract          = tract_labels[response],
      beta_std       = strip0(sprintf("%.3f", as.numeric(estimate))),
      statistic      = strip0(sprintf("%.3f", as.numeric(statistic))),
      `p (OLS)`      = strip0(sprintf("%.3f", as.numeric(p.value))),
      `pFDR (OLS)`   = strip0(sprintf("%.3f", as.numeric(.data[[p_fdr_col]]))),
      `p (HC3)`      = strip0(sprintf("%.3f", p_HC3)),
      `pFDR (HC3)`   = strip0(sprintf("%.3f", pFDR_HC3)),
      `95% CI (HC3)` = paste0("[", strip0(sprintf("%.3f", HC3_lo_std)), ", ",
                               strip0(sprintf("%.3f", HC3_hi_std)), "]"),
      `95% CI (BCa)` = paste0("[", strip0(sprintf("%.3f", BCa_lo_std)), ", ",
                               strip0(sprintf("%.3f", BCa_hi_std)), "]")
    ) |>
    dplyr::select(
      Tract, beta_std, statistic,
      `p (OLS)`, `pFDR (OLS)`,
      `p (HC3)`, `pFDR (HC3)`,
      `95% CI (HC3)`, `95% CI (BCa)`
    ) |>
    dplyr::arrange(factor(Tract, levels = tract_labels))
}


# make_gt() ────────────────────────────────────────────────────────────────────
make_gt <- function(tbl, caption, note) {
  tbl |>
    gt() |>
    tab_header(title = html(paste0("<b>", caption, "</b>"))) |>
    tab_style(
      style     = cell_text(align = "left"),
      locations = cells_title(groups = "title")
    ) |>
    cols_label(
      Tract          = "Tract",
      beta_std       = html("\u03b2"),
      statistic      = html("<i>t</i>"),
      `p (OLS)`      = html("<i>p</i> (OLS)"),
      `pFDR (OLS)`   = html("<i>p</i><sub>FDR</sub> (OLS)"),
      `p (HC3)`      = html("<i>p</i> (HC3)"),
      `pFDR (HC3)`   = html("<i>p</i><sub>FDR</sub> (HC3)"),
      `95% CI (HC3)` = "95% CI (HC3)",
      `95% CI (BCa)` = "95% CI (BCa)"
    ) |>
    cols_align(align = "left",   columns = Tract) |>
    cols_align(align = "center", columns = -Tract) |>
    tab_options(
      table.font.size                   = "small",
      source_notes.font.size            = "small",
      data_row.padding                  = px(2),
      heading.border.bottom.style       = "hidden",
      table.border.top.style            = "solid",
      table.border.bottom.style         = "solid",
      column_labels.border.top.style    = "solid",
      column_labels.border.bottom.style = "solid",
      table_body.border.bottom.style    = "solid"
    ) |>
    tab_source_note(source_note = html(paste0("<i>Note.</i> ", note)))
}


# ==============================================================================
# Helpers for Table S2
# ==============================================================================

# tract_labels / tract_prefix() ────────────────────────────────────────────────
# Map full WM column names to readable tract labels
tract_labels <- c(
  t_af_left      = "AF left",      t_af_right      = "AF right",
  t_cb_left      = "CB left",      t_cb_right      = "CB right",
  t_cc1          = "CC 1",         t_cc2           = "CC 2",
  t_cc3          = "CC 3",         t_cc4           = "CC 4",
  t_cc5          = "CC 5",         t_cc6           = "CC 6",
  t_cc7          = "CC 7",
  t_ilf_left     = "ILF left",     t_ilf_right     = "ILF right",
  t_ioff_left    = "IoFF left",    t_ioff_right    = "IoFF right",
  t_slf.i_left   = "SLF I left",   t_slf.i_right   = "SLF I right",
  t_slf.ii_left  = "SLF II left",  t_slf.ii_right  = "SLF II right",
  t_slf.iii_left = "SLF III left", t_slf.iii_right = "SLF III right",
  t_uf_left      = "UF left",      t_uf_right      = "UF right"
)

tract_prefix <- function(x) sub("\\.(fa1|tensor1\\..+)\\.mean$", "", x)


# fit_wm() ─────────────────────────────────────────────────────────────────────
# Fit a multivariate linear regression for one WM metric.
#
# Returns a tibble: tract, statistic, p_fdr, beta.
fit_wm <- function(dat,
                   col_pattern,
                   predictor   = "ctq_totalscore",
                   covars_cont = "age",
                   covars_cat  = "gender") {

  y_mat <- dat %>% select(matches(col_pattern)) %>% as.matrix()
  rhs   <- paste(c(predictor, covars_cont, covars_cat), collapse = " + ")

  # Unstandardised model: FDR across all terms × tracts, then filter to predictor
  res_unst <- lm(as.formula(paste("y_mat ~", rhs)), data = dat) %>%
    broom::tidy() %>%
    mutate(p_fdr = p.adjust(p.value, method = "fdr")) %>%
    filter(term == predictor) %>%
    mutate(tract = tract_labels[tract_prefix(response)]) %>%
    select(tract, statistic, p_fdr)

  # Standardised model: z-score continuous predictors, extract β
  dat_sc <- dat %>%
    mutate(across(all_of(c(predictor, covars_cont)), ~ as.numeric(scale(.x))))

  res_sc <- lm(as.formula(paste("scale(y_mat) ~", rhs)), data = dat_sc) %>%
    broom::tidy() %>%
    filter(term == predictor) %>%
    mutate(tract = tract_labels[tract_prefix(response)]) %>%
    select(tract, beta = estimate)

  left_join(res_unst, res_sc, by = "tract")
}


# make_s2_table() ──────────────────────────────────────────────────────────────
# Render an APA-style gt table for Table S2.
#
# `results`     : joined tibble (tract, FA_p_fdr, FA_beta, MD_*, AD_*, RD_*).
# `table_label` : bold table identifier, e.g. "Table S2a."
# `covar_note`  : covariate string for the source note.
# `n`           : sample size shown in the title.
make_s2_table <- function(results, table_label, covar_note, n) {
  results |>
    dplyr::select(tract,
                  FA_p_fdr, FA_beta,
                  MD_p_fdr, MD_beta,
                  AD_p_fdr, AD_beta,
                  RD_p_fdr, RD_beta) |>
    gt::gt() |>
    gt::tab_header(
      title = gt::html(paste0(
        "<b>", table_label, "</b> Multiple regression results between childhood ",
        "maltreatment severity and white matter microstructure (<i>N</i> = ", n, ")."
      ))
    ) |>
    gt::tab_style(
      style     = gt::cell_text(align = "left"),
      locations = gt::cells_title(groups = "title")
    ) |>
    gt::tab_spanner(label = gt::md("**Fractional anisotropy (FA)**"), columns = c(FA_p_fdr, FA_beta)) |>
    gt::tab_spanner(label = gt::md("**Mean diffusivity (MD)**"),      columns = c(MD_p_fdr, MD_beta)) |>
    gt::tab_spanner(label = gt::md("**Axial diffusivity (AD)**"),     columns = c(AD_p_fdr, AD_beta)) |>
    gt::tab_spanner(label = gt::md("**Radial diffusivity (RD)**"),    columns = c(RD_p_fdr, RD_beta)) |>
    gt::cols_label(
      tract    = "",
      FA_p_fdr = gt::html("<i>p</i><sub>FDR</sub>"), FA_beta = gt::md("**beta**"),
      MD_p_fdr = gt::html("<i>p</i><sub>FDR</sub>"), MD_beta = gt::md("**beta**"),
      AD_p_fdr = gt::html("<i>p</i><sub>FDR</sub>"), AD_beta = gt::md("**beta**"),
      RD_p_fdr = gt::html("<i>p</i><sub>FDR</sub>"), RD_beta = gt::md("**beta**")
    ) |>
    gt::fmt_number(columns = -tract, decimals = 3) |>
    gt::text_transform(
      locations = gt::cells_body(columns = -tract),
      fn = function(x) {
        x <- sub("^0\\.",       ".",       x)   # 0.xxx  -> .xxx
        x <- sub("^\u22120\\.", "\u2212.", x)   # −0.xxx -> −.xxx
        x
      }
    ) |>
    gt::tab_style(style = gt::cell_text(weight = "bold"),
                  locations = gt::cells_body(columns = c(FA_p_fdr, FA_beta), rows = FA_p_fdr < 0.05)) |>
    gt::tab_style(style = gt::cell_text(weight = "bold"),
                  locations = gt::cells_body(columns = c(MD_p_fdr, MD_beta), rows = MD_p_fdr < 0.05)) |>
    gt::tab_style(style = gt::cell_text(weight = "bold"),
                  locations = gt::cells_body(columns = c(AD_p_fdr, AD_beta), rows = AD_p_fdr < 0.05)) |>
    gt::tab_style(style = gt::cell_text(weight = "bold"),
                  locations = gt::cells_body(columns = c(RD_p_fdr, RD_beta), rows = RD_p_fdr < 0.05)) |>
    gt::tab_style(style = gt::cell_text(weight = "bold"),
                  locations = gt::cells_body(columns = tract)) |>
    gt::cols_align(align = "left",   columns = tract) |>
    gt::cols_align(align = "center", columns = -tract) |>
    gt::tab_options(
      table.font.size                   = "small",
      source_notes.font.size            = "small",
      data_row.padding                  = gt::px(2),
      heading.border.bottom.style       = "hidden",
      table.border.top.style            = "solid",
      table.border.bottom.style         = "solid",
      column_labels.border.top.style    = "solid",
      column_labels.border.bottom.style = "solid",
      table_body.border.bottom.style    = "solid"
    ) |>
    gt::tab_source_note(
      source_note = gt::html(paste0(
        "All analyses were corrected for ", covar_note, " as covariates.<br>",
        "<i>Note.</i> FA, fractional anisotropy; MD, mean diffusivity; AD, axial diffusivity; ",
        "RD, radial diffusivity; FDR, false-discovery rate; AF, arcuate fasciculus; ",
        "CB, cingulum bundle; CC, corpus callosum; ILF, inferior longitudinal fasciculus; ",
        "IoFF, inferior occipito-frontal fasciculus; SLF, superior longitudinal fasciculus; ",
        "UF, uncinate fasciculus."
      ))
    )
}
```

---

# Statistical Analysis

> *Overview:* This document contains four supplementary analyses extending the
> primary H1a regression results (*N* = `r nrow(dat)` participants, 23 WM tracts,
> four diffusion metrics: FA, MD, AD, RD). Unless noted otherwise, all analyses
> share the same predictor (**CTQ total score**), covariates (**age, gender**), and
> FDR correction strategy (**Benjamini–Hochberg, *q* < .05, applied within each
> diffusion metric across all model terms × 23 tracts**) as the primary analysis.

---

# Table S2: Post-Hoc Analyses with Additional Covariates

> *Methods (Table S2):* The primary H1a regressions are repeated twice with
> expanded covariate sets to assess whether associations between CTQ and WM
> microstructure are robust to demographic confounds not included in the primary
> model. **Model 1** adds race. **Model 2** adds race and alcohol use.
>
> - **Outcome (Y):** WM tract metric (FA, MD, AD, RD), 23 tracts
> - **Predictor (X):** CTQ total score
> - **Model 1 covariates:** Age, gender, race
> - **Model 2 covariates:** Age, gender, race, alcohol use (AUDIT-10)

The same `fit_wm()` function used in the primary H1a analysis is used here
with the `covars_cont` argument extended to include `race` (Model 1) and
additionally `audit_10_score` (Model 2). Continuous additional covariates
are z-scored in the standardised model alongside age and CTQ. Results are
displayed with the same gt table style as Table 2 in the main script.

---

## Model 1: Age, Gender, Race

```{r s2_model1, echo=FALSE}
# Column patterns for each metric (same as primary H1a analysis)
metrics <- list(
  FA = "\\.fa1\\.mean$",
  MD = "\\.tensor1\\.meandiffusivity\\.mean$",
  AD = "\\.tensor1\\.maxeigenvalue\\.mean$",
  RD = "\\.tensor1\\.mineigenvalue\\.mean$"
)

# Run all four metrics with race added as a continuous covariate, then join
results_m1 <- imap(metrics, function(pattern, metric) {
  fit_wm(dat, pattern,
         covars_cont = c("age", "race"),
         covars_cat  = "gender") |>
    rename_with(~ paste0(metric, "_", .x), .cols = -tract)
}) |>
  reduce(left_join, by = "tract")

tbl_s2a <- make_s2_table(
  results_m1,
  table_label = "Table S2a.",
  covar_note  = "age, gender, and race",
  n           = nrow(dat)
)

tbl_s2a

gtsave(tbl_s2a,
       here("output", "tables", "supplementary", "TableS2a_CTQ_WM_Race.html"))
```

---

## Model 2: Age, Gender, Race, Alcohol Use

```{r s2_model2, echo=FALSE}
# Run all four metrics with race and alcohol use added as continuous covariates
results_m2 <- imap(metrics, function(pattern, metric) {
  fit_wm(dat, pattern,
         covars_cont = c("age", "race", "audit_10_score"),
         covars_cat  = "gender") |>
    rename_with(~ paste0(metric, "_", .x), .cols = -tract)
}) |>
  reduce(left_join, by = "tract")

tbl_s2b <- make_s2_table(
  results_m2,
  table_label = "Table S2b.",
  covar_note  = "age, gender, race, and alcohol use (AUDIT-10)",
  n           = nrow(dat)
)

tbl_s2b

gtsave(tbl_s2b,
       here("output", "tables", "supplementary", "TableS2b_CTQ_WM_RaceAlcohol.html"))
```

---

# Table S5: Robustness to Heteroscedasticity — HC3 SEs and BCa Bootstrap CIs

> *Methods (Table S5):* To evaluate whether the primary H1a results are robust
> to heteroscedasticity, all regressions were re-estimated using
> heteroskedasticity-consistent (HC3) standard errors (MacKinnon & White, 1985)
> and bias-corrected accelerated (BCa) bootstrap 95% confidence intervals
> (5,000 resamples, seed = 1234).
>
> For each WM metric, the table reports:
>
> - **beta (std):** standardised OLS regression coefficient for CTQ
> - **t:** *t*-statistic from the unstandardised OLS model
> - **p (OLS) / pFDR (OLS):** raw and FDR-corrected *p*-values from the standard OLS model
> - **p (HC3) / pFDR (HC3):** raw and FDR-corrected *p*-values using HC3 standard errors
> - **95% CI (HC3):** HC3-based confidence interval on the standardised beta
> - **95% CI (BCa):** bias-corrected accelerated bootstrap CI on the standardised beta
>
> FDR correction (Benjamini–Hochberg) is applied separately within each
> diffusion metric across **all model terms × 23 tracts**.
>
> - **Outcome (Y):** WM tract metric (FA, MD, AD, RD), 23 tracts each
> - **Predictor (X):** CTQ total score
> - **Covariates:** Age, gender
> - **Output files:** `figs/df_regression_{fa,md,ad,rd}_hc3.bca.docx`

---

## Fractional Anisotropy (FA)

```{r hc3_fa, echo=FALSE}
fa_cols   <- paste0(tract_stems, metric_suffixes[["FA"]])
fa_values <- as.matrix(dat[, fa_cols])
colnames(fa_values) <- NULL          # broom::tidy() will label responses Y1..Y23
fa_names  <- paste0("Y", seq_len(ncol(fa_values)))

fa_ols     <- fit_ols_wm(fa_values, dat, "ctq_totalscore", "p.fdr.fa")
lm_fa_out  <- fa_ols$out
lm_fa_beta <- fa_ols$beta

hc3_fa_std_ctq <- fit_hc3_bca(fa_values, fa_names, dat,
                               predictor = "ctq_totalscore",
                               R_boot    = 5000)

fa_table_std <- build_hc3_table(lm_fa_out, lm_fa_beta, hc3_fa_std_ctq,
                                tract_pretty, "p.fdr.fa")

hc3_note <- paste0(
  "OLS = ordinary least squares; HC3 = heteroskedasticity-consistent standard errors; ",
  "BCa = bias-corrected and accelerated bootstrap; FDR = false-discovery rate; ",
  "\u03b2 = standardised regression coefficient; CI = confidence interval. ",
  "All models adjusted for age and gender. FDR correction applied across all model ",
  "terms \u00d7 tracts."
)

tbl_fa <- make_gt(
  fa_table_std,
  caption = "Supplementary Table S5a. Multiple regression results for FA: CTQ and white matter microstructure with HC3 standard errors and BCa confidence intervals.",
  note    = hc3_note
)
gtsave(tbl_fa, here("output", "tables", "supplementary", "TableS5a_HC3_FA.html"))
tbl_fa
```

---

## Mean Diffusivity (MD)

```{r hc3_md, echo=FALSE}
md_cols   <- paste0(tract_stems, metric_suffixes[["MD"]])
md_values <- as.matrix(dat[, md_cols])
colnames(md_values) <- NULL
md_names  <- paste0("Y", seq_len(ncol(md_values)))

md_ols     <- fit_ols_wm(md_values, dat, "ctq_totalscore", "p.fdr.md")
lm_md_out  <- md_ols$out
lm_md_beta <- md_ols$beta

hc3_md_std_ctq <- fit_hc3_bca(md_values, md_names, dat,
                               predictor = "ctq_totalscore",
                               R_boot    = 5000)

md_table_std <- build_hc3_table(lm_md_out, lm_md_beta, hc3_md_std_ctq,
                                tract_pretty, "p.fdr.md")

tbl_md <- make_gt(
  md_table_std,
  caption = "Supplementary Table S5b. Multiple regression results for MD: CTQ and white matter microstructure with HC3 standard errors and BCa confidence intervals.",
  note    = hc3_note
)
gtsave(tbl_md, here("output", "tables", "supplementary", "TableS5b_HC3_MD.html"))
tbl_md
```

---

## Axial Diffusivity (AD)

```{r hc3_ad, echo=FALSE}
ad_cols   <- paste0(tract_stems, metric_suffixes[["AD"]])
ad_values <- as.matrix(dat[, ad_cols])
colnames(ad_values) <- NULL
ad_names  <- paste0("Y", seq_len(ncol(ad_values)))

ad_ols     <- fit_ols_wm(ad_values, dat, "ctq_totalscore", "p.fdr.ad")
lm_ad_out  <- ad_ols$out
lm_ad_beta <- ad_ols$beta

hc3_ad_std_ctq <- fit_hc3_bca(ad_values, ad_names, dat,
                               predictor = "ctq_totalscore",
                               R_boot    = 5000)

ad_table_std <- build_hc3_table(lm_ad_out, lm_ad_beta, hc3_ad_std_ctq,
                                tract_pretty, "p.fdr.ad")

tbl_ad <- make_gt(
  ad_table_std,
  caption = "Supplementary Table S5c. Multiple regression results for AD: CTQ and white matter microstructure with HC3 standard errors and BCa confidence intervals.",
  note    = hc3_note
)
gtsave(tbl_ad, here("output", "tables", "supplementary", "TableS5c_HC3_AD.html"))
tbl_ad
```

---

## Radial Diffusivity (RD)

```{r hc3_rd, echo=FALSE}
rd_cols   <- paste0(tract_stems, metric_suffixes[["RD"]])
rd_values <- as.matrix(dat[, rd_cols])
colnames(rd_values) <- NULL
rd_names  <- paste0("Y", seq_len(ncol(rd_values)))

rd_ols     <- fit_ols_wm(rd_values, dat, "ctq_totalscore", "p.fdr.rd")
lm_rd_out  <- rd_ols$out
lm_rd_beta <- rd_ols$beta

hc3_rd_std_ctq <- fit_hc3_bca(rd_values, rd_names, dat,
                               predictor = "ctq_totalscore",
                               R_boot    = 5000)

rd_table_std <- build_hc3_table(lm_rd_out, lm_rd_beta, hc3_rd_std_ctq,
                                tract_pretty, "p.fdr.rd")

tbl_rd <- make_gt(
  rd_table_std,
  caption = "Supplementary Table S5d. Multiple regression results for RD: CTQ and white matter microstructure with HC3 standard errors and BCa confidence intervals.",
  note    = hc3_note
)
gtsave(tbl_rd, here("output", "tables", "supplementary", "TableS5d_HC3_RD.html"))
tbl_rd
```

---

# Table S6: Sensitivity Analysis — Excluding PTSD-Only Subgroup

```{r noptsd_setup, include=FALSE}
# Exclude participants whose only diagnosis is PTSD (no mTBI or HC).
# This checks whether the small PTSD-only subgroup unduly influences results.
df_no.ptsd <- subset(dat, hc.ptsd.mtbi.groups.4groups != "PTSD-only")

# Re-state type coercions explicitly so this chunk is self-contained.
df_no.ptsd$ctq_totalscore <- as.numeric(df_no.ptsd$ctq_totalscore)
df_no.ptsd$age             <- as.numeric(df_no.ptsd$age)
df_no.ptsd$gender          <- as.numeric(df_no.ptsd$gender)
```

> *Methods (Table S6):* The primary H1a regressions are repeated after excluding
> the PTSD-only subgroup (*n* = `r nrow(dat) - nrow(df_no.ptsd)`) to verify that
> this  group does not disproportionately drive the main findings.
>
> - **Outcome (Y):** WM tract metric (FA, MD, AD, RD), 23 tracts each
> - **Predictor (X):** CTQ total score
> - **Covariates:** Age, gender
> - **Exclusion criterion:** `hc.ptsd.mtbi.groups.4groups == "PTSD-only"`
> - **Output file:** `figs/df_no_ptsd_table.docx`

---

## FA

```{r noptsd_fa, echo=FALSE}
fa_values <- as.matrix(df_no.ptsd[, paste0(tract_stems, metric_suffixes[["FA"]])])
colnames(fa_values) <- NULL

fa_ols       <- fit_ols_wm(fa_values, df_no.ptsd, "ctq_totalscore", "p.fdr.fa")
lm_fa        <- fa_ols$out
lm_fa_scaled <- fa_ols$beta
```

## MD

```{r noptsd_md, echo=FALSE}
md_values <- as.matrix(df_no.ptsd[, paste0(tract_stems, metric_suffixes[["MD"]])])
colnames(md_values) <- NULL

md_ols       <- fit_ols_wm(md_values, df_no.ptsd, "ctq_totalscore", "p.fdr.md")
lm_md        <- md_ols$out
lm_md_scaled <- md_ols$beta
```

## AD

```{r noptsd_ad, echo=FALSE}
ad_values <- as.matrix(df_no.ptsd[, paste0(tract_stems, metric_suffixes[["AD"]])])
colnames(ad_values) <- NULL

ad_ols       <- fit_ols_wm(ad_values, df_no.ptsd, "ctq_totalscore", "p.fdr.ad")
lm_ad        <- ad_ols$out
lm_ad_scaled <- ad_ols$beta
```

## RD

```{r noptsd_rd, echo=FALSE}
rd_values <- as.matrix(df_no.ptsd[, paste0(tract_stems, metric_suffixes[["RD"]])])
colnames(rd_values) <- NULL

rd_ols       <- fit_ols_wm(rd_values, df_no.ptsd, "ctq_totalscore", "p.fdr.rd")
lm_rd        <- rd_ols$out
lm_rd_scaled <- rd_ols$beta
```

## Results Table (Table S6)

```{r noptsd_merge_p, echo=FALSE}
# Join FDR-corrected p-values across metrics
lm_fa_sel <- lm_fa |> dplyr::select(response, p.fdr.fa)
lm_md_sel <- lm_md |> dplyr::select(response, p.fdr.md)
lm_ad_sel <- lm_ad |> dplyr::select(response, p.fdr.ad)
lm_rd_sel <- lm_rd |> dplyr::select(response, p.fdr.rd)

df_combined <- full_join(lm_fa_sel, lm_md_sel, by = "response") |>
  full_join(lm_ad_sel, by = "response") |>
  full_join(lm_rd_sel, by = "response") |>
  dplyr::mutate(response = tract_pretty[response])

names(df_combined) <- c("White matter tracts",
                        "FA p-value (FDR-corrected)",
                        "MD p-value (FDR-corrected)",
                        "AD p-value (FDR-corrected)",
                        "RD p-value (FDR-corrected)")
```

```{r noptsd_merge_beta, echo=FALSE}
df_combined_scaled <- full_join(lm_fa_scaled, lm_md_scaled, by = "response") |>
  full_join(lm_ad_scaled, by = "response") |>
  full_join(lm_rd_scaled, by = "response") |>
  dplyr::mutate(response = tract_pretty[response])

names(df_combined_scaled) <- c("White matter tracts",
                               "FA beta", "MD beta", "AD beta", "RD beta")

# Combine FDR-corrected p-values and standardised betas; order columns by metric
df_combined_full <- full_join(df_combined, df_combined_scaled,
                              by = "White matter tracts") |>
  dplyr::select(contains("White matter tracts"),
                contains("FA"), contains("MD"),
                contains("AD"), contains("RD"))

tbl_s6 <- df_combined_full |>
  gt() |>
  tab_header(
    title = html(paste0(
      "<b>Supplementary Table S6.</b> Multiple regression results between CTQ total score ",
      "and white matter microstructure, excluding PTSD-only participants (N = ",
      nrow(df_no.ptsd), ")."
    ))
  ) |>
  tab_style(
    style     = cell_text(align = "left"),
    locations = cells_title(groups = "title")
  ) |>
  tab_spanner(label = md("**FA**"), columns = c(`FA p-value (FDR-corrected)`, `FA beta`)) |>
  tab_spanner(label = md("**MD**"), columns = c(`MD p-value (FDR-corrected)`, `MD beta`)) |>
  tab_spanner(label = md("**AD**"), columns = c(`AD p-value (FDR-corrected)`, `AD beta`)) |>
  tab_spanner(label = md("**RD**"), columns = c(`RD p-value (FDR-corrected)`, `RD beta`)) |>
  cols_label(
    `White matter tracts`        = "",
    `FA p-value (FDR-corrected)` = html("<i>p</i><sub>FDR</sub>"),
    `FA beta`                    = html("\u03b2"),
    `MD p-value (FDR-corrected)` = html("<i>p</i><sub>FDR</sub>"),
    `MD beta`                    = html("\u03b2"),
    `AD p-value (FDR-corrected)` = html("<i>p</i><sub>FDR</sub>"),
    `AD beta`                    = html("\u03b2"),
    `RD p-value (FDR-corrected)` = html("<i>p</i><sub>FDR</sub>"),
    `RD beta`                    = html("\u03b2")
  ) |>
  fmt_number(columns = c(`FA beta`, `MD beta`, `AD beta`, `RD beta`), decimals = 3) |>
  text_transform(
    locations = cells_body(columns = c(`FA beta`, `MD beta`, `AD beta`, `RD beta`)),
    fn = function(x) {
      x <- sub("^0\\.",       ".",       x)   # 0.xxx  -> .xxx
      x <- sub("^\u22120\\.", "\u2212.", x)   # −0.xxx -> −.xxx
      x
    }
  ) |>
  fmt_missing(columns = everything(), missing_text = "\u2014") |>
  cols_align(align = "left",   columns = `White matter tracts`) |>
  cols_align(align = "center", columns = -`White matter tracts`) |>
  tab_options(
    table.font.size                   = "small",
    source_notes.font.size            = "small",
    data_row.padding                  = px(2),
    heading.border.bottom.style       = "hidden",
    table.border.top.style            = "solid",
    table.border.bottom.style         = "solid",
    column_labels.border.top.style    = "solid",
    column_labels.border.bottom.style = "solid",
    table_body.border.bottom.style    = "solid"
  ) |>
  tab_source_note(source_note = html(paste0(
    "<i>Note.</i> Higher CTQ scores were significantly associated with lower mean diffusivity ",
    "and axial diffusivity in various white matter tracts. No significant associations were ",
    "found for fractional anisotropy or radial diffusivity. All analyses were adjusted for ",
    "age and gender. FA, fractional anisotropy; MD, mean diffusivity; AD, axial diffusivity; ",
    "RD, radial diffusivity; FDR, false-discovery rate; AF, arcuate fasciculus; CB, cingulum ",
    "bundle; CC, corpus callosum; ILF, inferior longitudinal fasciculus; IoFF, inferior ",
    "occipito-frontal fasciculus; SLF, superior longitudinal fasciculus; UF, uncinate fasciculus."
  )))

gtsave(tbl_s6, here("output", "tables", "supplementary", "TableS6_Sensitivity_NoPTSD.html"))
tbl_s6
```

---

# Table S12: CTQ Subscale Analyses — MD and AD

> *Methods (Table S12):* To identify which aspects of childhood maltreatment drive
> the primary MD and AD associations, analyses were repeated with each CTQ subscale
> as the predictor in place of the total score.
>
> - **Outcome (Y):** MD tracts (significant in primary analysis) and AD tracts (significant in primary analysis)
> - **Predictors (X):** Five CTQ subscales — emotional neglect, physical neglect, emotional abuse, physical abuse, sexual abuse
> - **Covariates:** Age, gender
> - **FDR correction:** Benjamini–Hochberg within each metric × subscale combination

---

## Analysis

```{r s12_analysis, echo=FALSE}
# CTQ subscales: name = display label, value = column in dat
subscales <- list(
  "Emotional neglect" = "ctq_emoneg_score",
  "Physical neglect"  = "ctq_physneg_score",
  "Emotional abuse"   = "ctq_emo_score",
  "Physical abuse"    = "ctq_phys_score",
  "Sexual abuse"      = "ctq_sex_score"
)

# Short prefixes for column naming (order matches subscales list)
subscale_keys <- c("en", "pn", "ea", "pa", "sa")

md_pattern <- "\\.tensor1\\.meandiffusivity\\.mean$"
ad_pattern <- "\\.tensor1\\.maxeigenvalue\\.mean$"

# Significant-tract subsets (from primary analysis), fixed order as in Table S12
sig_md_tracts <- c(
  "AF right", "CC 1", "CC 2", "CC 6",
  "ILF right", "IoFF left", "IoFF right", "UF right"
)

sig_ad_tracts <- c(
  "AF left", "AF right", "CB right",
  "CC 2", "CC 3", "CC 4", "CC 5", "CC 6",
  "IoFF left", "SLF I left", "SLF I right",
  "SLF II left", "SLF II right", "SLF III right", "UF right"
)

# run_subscale_wide(): Run fit_wm() for all 5 subscales on one metric, return
# wide tibble (tract + 2 cols per subscale: <key>_pfdr, <key>_beta).
# Filtered to sig_tracts before returning.
run_subscale_wide <- function(metric_pattern, sig_tracts) {
  imap(subscales, function(col, label) {
    key <- subscale_keys[match(label, names(subscales))]
    fit_wm(dat, metric_pattern, predictor = col) %>%
      select(tract, p_fdr, beta) %>%
      rename(!!paste0(key, "_pfdr") := p_fdr,
             !!paste0(key, "_beta") := beta)
  }) %>%
    reduce(left_join, by = "tract") %>%
    filter(tract %in% sig_tracts) %>%
    slice(match(sig_tracts, tract))
}

md_wide <- run_subscale_wide(md_pattern, sig_md_tracts) %>%
  mutate(group = "MD", .before = tract)
ad_wide <- run_subscale_wide(ad_pattern, sig_ad_tracts) %>%
  mutate(group = "AD", .before = tract)

results_s12 <- bind_rows(md_wide, ad_wide)
```

## Results Table (Table S12)

```{r s12_table, echo=FALSE}
pfdr_cols <- paste0(subscale_keys, "_pfdr")
beta_cols  <- paste0(subscale_keys, "_beta")
num_cols   <- c(pfdr_cols, beta_cols)

# Capture bold row indices from numeric data BEFORE character conversion
bold_idx <- lapply(subscale_keys, function(k) which(results_s12[[paste0(k, "_pfdr")]] < 0.05))
names(bold_idx) <- subscale_keys

# Format numeric columns to 3-decimal character strings with leading-zero stripped
fmt3 <- function(x) {
  s <- formatC(x, digits = 3, format = "f")
  s <- sub("^0\\.",   ".",  s)
  s <- sub("^-0\\.", "-.", s)
  s
}

results_display <- results_s12 %>%
  mutate(across(all_of(num_cols), fmt3))

tbl_s12 <- results_display |>
  gt(groupname_col = "group") |>
  tab_header(
    title = html(paste0(
      "<b>Supplementary Table S12.</b> Multiple regression results using childhood maltreatment ",
      "subtype as the criterion (N = ", nrow(dat), ")."
    ))
  ) |>
  tab_style(
    style     = cell_text(align = "left"),
    locations = cells_title(groups = "title")
  ) |>
  tab_style(
    style     = cell_text(weight = "bold"),
    locations = cells_row_groups()
  ) |>
  # ── Subscale column spanners ──────────────────────────────────────────────
  tab_spanner(label = md(paste0("**", names(subscales)[1], "**")),
              columns = c(en_pfdr, en_beta)) |>
  tab_spanner(label = md(paste0("**", names(subscales)[2], "**")),
              columns = c(pn_pfdr, pn_beta)) |>
  tab_spanner(label = md(paste0("**", names(subscales)[3], "**")),
              columns = c(ea_pfdr, ea_beta)) |>
  tab_spanner(label = md(paste0("**", names(subscales)[4], "**")),
              columns = c(pa_pfdr, pa_beta)) |>
  tab_spanner(label = md(paste0("**", names(subscales)[5], "**")),
              columns = c(sa_pfdr, sa_beta)) |>
  # ── Sub-column labels ─────────────────────────────────────────────────────
  cols_label(
    tract   = "",
    en_pfdr = html("<i>p</i><sub>FDR</sub>"), en_beta = html("\u03b2"),
    pn_pfdr = html("<i>p</i><sub>FDR</sub>"), pn_beta = html("\u03b2"),
    ea_pfdr = html("<i>p</i><sub>FDR</sub>"), ea_beta = html("\u03b2"),
    pa_pfdr = html("<i>p</i><sub>FDR</sub>"), pa_beta = html("\u03b2"),
    sa_pfdr = html("<i>p</i><sub>FDR</sub>"), sa_beta = html("\u03b2")
  ) |>
  # ── Bold significant pFDR + companion beta ────────────────────────────────
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_body(columns = c(en_pfdr, en_beta),
                                   rows    = bold_idx[["en"]])) |>
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_body(columns = c(pn_pfdr, pn_beta),
                                   rows    = bold_idx[["pn"]])) |>
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_body(columns = c(ea_pfdr, ea_beta),
                                   rows    = bold_idx[["ea"]])) |>
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_body(columns = c(pa_pfdr, pa_beta),
                                   rows    = bold_idx[["pa"]])) |>
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_body(columns = c(sa_pfdr, sa_beta),
                                   rows    = bold_idx[["sa"]])) |>
  # ── Alignment ─────────────────────────────────────────────────────────────
  cols_align(align = "left",   columns = tract) |>
  cols_align(align = "center", columns = all_of(num_cols)) |>
  # ── APA options ───────────────────────────────────────────────────────────
  tab_options(
    table.font.size                   = "small",
    source_notes.font.size            = "small",
    data_row.padding                  = px(2),
    heading.border.bottom.style       = "hidden",
    table.border.top.style            = "solid",
    table.border.bottom.style         = "solid",
    column_labels.border.top.style    = "solid",
    column_labels.border.bottom.style = "solid",
    table_body.border.bottom.style    = "solid",
    row_group.border.top.style        = "hidden",
    row_group.border.bottom.style     = "hidden"
  ) |>
  tab_source_note(source_note = html(paste0(
    "<i>Note.</i> CTQ, Childhood Trauma Questionnaire; FDR, false-discovery rate; ",
    "MD, mean diffusivity; AD, axial diffusivity; AF, arcuate fasciculus; CB, cingulum ",
    "bundle; CC, corpus callosum; ILF, inferior longitudinal fasciculus; IoFF, inferior ",
    "occipito-frontal fasciculus; SLF, superior longitudinal fasciculus; UF, uncinate ",
    "fasciculus. Additional regression analyses using CTQ subscales as independent variables ",
    "highlighted the substantial contributions of emotional neglect and emotional abuse to ",
    "the observed effects. All analyses were corrected for age and gender."
  )))

gtsave(tbl_s12, here("output", "tables", "supplementary", "TableS12_CTQ_Subscales.html"))
tbl_s12
```
