---
title: "INTRuST CM — H1b Supplementary: Robustness of Moderation Analyses"
author: "Luisa Berger"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

suppressPackageStartupMessages({
  library(here)
  library(tidyverse)
  library(fastDummies)
  library(sandwich)
  library(lmtest)
  library(gt)
  library(boot)
  library(glue)
})

set.seed(1234)

dat <- readr::read_csv(
  here("data", "analysis_simulated.csv"),
  show_col_types = FALSE,
  name_repair = "minimal"
)

dat$gender     <- as.numeric(dat$gender)
dat$tbi.diagnosis  <- as.numeric(dat$tbi.diagnosis)
dat$ptsd.diagnosis <- as.numeric(dat$ptsd.diagnosis)
dat$race       <- as.numeric(dat$race)

# Dummy-code the 4-group diagnosis variable (reference group: HC)
dat <- dummy_cols(dat, select_columns = "hc.ptsd.mtbi.groups.4groups")
dat$hc.ptsd.mtbi.groups.4groups                        <- as.numeric(dat$hc.ptsd.mtbi.groups.4groups)
dat$`hc.ptsd.mtbi.groups.4groups_mTBI-only`            <- as.numeric(dat$`hc.ptsd.mtbi.groups.4groups_mTBI-only`)
dat$`hc.ptsd.mtbi.groups.4groups_PTSD-only`            <- as.numeric(dat$`hc.ptsd.mtbi.groups.4groups_PTSD-only`)
dat$`hc.ptsd.mtbi.groups.4groups_Comorbid mTBI+PTSD`   <- as.numeric(dat$`hc.ptsd.mtbi.groups.4groups_Comorbid mTBI+PTSD`)
```

```{r define_settings, include=FALSE}
X   <- "ctq_totalscore"
COV <- c("age", "gender")

# Named vector: column name in dat -> readable label used in output tables
mod_map <- c(
  "hc.ptsd.mtbi.groups.4groups_PTSD-only"          = "PTSD",
  "hc.ptsd.mtbi.groups.4groups_mTBI-only"          = "mTBI",
  "hc.ptsd.mtbi.groups.4groups_Comorbid mTBI+PTSD" = "PTSD+mTBI"
)

# Tracts significant in H1a (FDR p < .05), ordered by diffusion metric and tract
tracts <- c(
  # Mean diffusivity (MD)
  "t_af_right.tensor1.meandiffusivity.mean",
  "t_ilf_right.tensor1.meandiffusivity.mean",
  "t_ioff_left.tensor1.meandiffusivity.mean",
  "t_ioff_right.tensor1.meandiffusivity.mean",
  "t_uf_right.tensor1.meandiffusivity.mean",
  "t_cc1.tensor1.meandiffusivity.mean",
  "t_cc2.tensor1.meandiffusivity.mean",
  "t_cc6.tensor1.meandiffusivity.mean",
  # Axial diffusivity (AD)
  "t_af_left.tensor1.maxeigenvalue.mean",
  "t_af_right.tensor1.maxeigenvalue.mean",
  "t_cb_right.tensor1.maxeigenvalue.mean",
  "t_ioff_left.tensor1.maxeigenvalue.mean",
  "t_slf.i_left.tensor1.maxeigenvalue.mean",
  "t_slf.i_right.tensor1.maxeigenvalue.mean",
  "t_slf.ii_left.tensor1.maxeigenvalue.mean",
  "t_slf.ii_right.tensor1.maxeigenvalue.mean",
  "t_slf.iii_right.tensor1.maxeigenvalue.mean",
  "t_cc2.tensor1.maxeigenvalue.mean",
  "t_cc3.tensor1.maxeigenvalue.mean",
  "t_cc4.tensor1.maxeigenvalue.mean",
  "t_cc5.tensor1.maxeigenvalue.mean",
  "t_cc6.tensor1.maxeigenvalue.mean"
)

# Maps raw column names to readable labels and display order within each metric group
tract_map <- tibble::tribble(
  ~Tract_raw,                                           ~Group, ~Tract_pretty,    ~Tract_order,
  # MD
  "t_af_right.tensor1.meandiffusivity.mean",            "MD",   "AF right",        1,
  "t_cc1.tensor1.meandiffusivity.mean",                 "MD",   "CC 1",            2,
  "t_cc2.tensor1.meandiffusivity.mean",                 "MD",   "CC 2",            3,
  "t_cc6.tensor1.meandiffusivity.mean",                 "MD",   "CC 6",            4,
  "t_ilf_right.tensor1.meandiffusivity.mean",           "MD",   "ILF right",       5,
  "t_ioff_left.tensor1.meandiffusivity.mean",           "MD",   "IoFF left",       6,
  "t_ioff_right.tensor1.meandiffusivity.mean",          "MD",   "IoFF right",      7,
  "t_uf_right.tensor1.meandiffusivity.mean",            "MD",   "UF right",        8,
  # AD
  "t_af_left.tensor1.maxeigenvalue.mean",               "AD",   "AF left",         1,
  "t_af_right.tensor1.maxeigenvalue.mean",              "AD",   "AF right",        2,
  "t_cb_right.tensor1.maxeigenvalue.mean",              "AD",   "CB right",        3,
  "t_cc2.tensor1.maxeigenvalue.mean",                   "AD",   "CC 2",            4,
  "t_cc3.tensor1.maxeigenvalue.mean",                   "AD",   "CC 3",            5,
  "t_cc4.tensor1.maxeigenvalue.mean",                   "AD",   "CC 4",            6,
  "t_cc5.tensor1.maxeigenvalue.mean",                   "AD",   "CC 5",            7,
  "t_cc6.tensor1.maxeigenvalue.mean",                   "AD",   "CC 6",            8,
  "t_ioff_left.tensor1.maxeigenvalue.mean",             "AD",   "IoFF left",       9,
  "t_slf.i_left.tensor1.maxeigenvalue.mean",            "AD",   "SLF I left",     10,
  "t_slf.i_right.tensor1.maxeigenvalue.mean",           "AD",   "SLF I right",    11,
  "t_slf.ii_left.tensor1.maxeigenvalue.mean",           "AD",   "SLF II left",    12,
  "t_slf.ii_right.tensor1.maxeigenvalue.mean",          "AD",   "SLF II right",   13,
  "t_slf.iii_right.tensor1.maxeigenvalue.mean",         "AD",   "SLF III right",  14
)

BOOT_R <- 5000   # number of bootstrap resamples (used identically in S7 and S8)

# Check that expected tracts and moderators are present in the data
available_tracts <- intersect(tracts, names(dat))
dropped_tracts   <- setdiff(tracts, names(dat))
available_mods   <- intersect(names(mod_map), names(dat))
dropped_mods     <- setdiff(names(mod_map), names(dat))

if (length(dropped_tracts)) message("Skipping missing tracts: ",    paste(dropped_tracts, collapse = ", "))
if (length(dropped_mods))   message("Skipping missing moderators: ", paste(dropped_mods,  collapse = ", "))
```

```{r helpers, include=FALSE}
# APA-style formatting helpers ─────────────────────────────────────────────────
# All formatted strings use 3 decimal places and no leading zero.

bt      <- function(x) paste0("`", x, "`")   # backtick-quote a column name for formula use
strip0  <- function(s) gsub("(?<=^|-)0\\.", ".", s, perl = TRUE)
fmt3    <- function(x) strip0(sprintf("%.3f", round(as.numeric(x), 3)))
fmt_b   <- fmt3   # alias for unstandardised b (same precision and APA convention)
apa_p   <- function(p) sub("^= 0\\.", "= .", ifelse(p < .001, "< .001", sprintf("= %.3f", round(p, 3))))
fmt_ci  <- function(lo, hi) paste0("[", fmt3(lo), ", ", fmt3(hi), "]")
fmt_r2  <- function(x) sub("^0", "", sprintf("%.3f", round(x, 3)))
fmt_pr2 <- function(x) fmt3(pmax(0, pmin(1, x)))

# get_ixname() ─────────────────────────────────────────────────────────────────
# Returns the coefficient name for the X:W interaction term in a fitted model.

get_ixname <- function(fit, X, W) {
  rn_clean <- gsub("`", "", names(coef(fit)), fixed = TRUE)
  cand     <- c(paste0(X, ":", W), paste0(W, ":", X))
  ix_clean <- intersect(cand, rn_clean)
  if (!length(ix_clean))
    stop(glue("Interaction {X}:{W} not found in model. Check variable names."))
  names(coef(fit))[match(ix_clean[1], rn_clean)]
}

# run_one(): CTQ × moderator interaction model (HC3 SEs + BCa CI); returns one-row tibble.
run_one <- function(y, w, data) {
  needed       <- c(y, X, w, COV)
  missing_cols <- setdiff(needed, names(data))
  if (length(missing_cols))
    stop(glue("run_one: missing columns for {y} x {w}: {paste(missing_cols, collapse = ', ')}"))

  fml <- as.formula(glue(
    "{bt(y)} ~ {bt(X)} * {bt(w)}{if (length(COV)) paste0(' + ', paste(bt(COV), collapse = ' + '))}"
  ))

  mf <- model.frame(fml, data = dat)

  # Standardise outcome, X, and numeric covariates; mean-centre moderator
  mfz      <- mf
  mfz[[y]] <- as.numeric(scale(mfz[[y]]))
  mfz[[X]] <- as.numeric(scale(mfz[[X]]))
  for (cv in COV) if (is.numeric(mfz[[cv]])) mfz[[cv]] <- as.numeric(scale(mfz[[cv]]))
  mw <- suppressWarnings(as.numeric(mfz[[w]]))
  if (all(is.finite(mw))) mfz[[w]] <- mw - mean(mw, na.rm = TRUE)

  fit <- lm(fml, data = mfz)
  ix  <- get_ixname(fit, X, w)

  # HC3 robust inference
  V      <- sandwich::vcovHC(fit, type = "HC3")
  est    <- unname(coef(fit)[ix])
  se_hc3 <- sqrt(diag(V))[ix]
  t_hc3  <- est / se_hc3
  p_hc3  <- 2 * pt(-abs(t_hc3), df = df.residual(fit))
  ci_hc3 <- lmtest::coefci(fit, vcov. = V)[ix, ]

  # Partial R² from the HC3 Wald statistic (1 df)
  pr2 <- as.numeric((t_hc3^2) / (t_hc3^2 + df.residual(fit)))

  # BCa bootstrap CI for the interaction coefficient
  stat_ix <- function(d, idx) {
    f <- try(lm(fml, data = d[idx, , drop = FALSE]), silent = TRUE)
    if (inherits(f, "try-error")) return(NA_real_)
    cf <- try(coef(f)[ix], silent = TRUE)
    if (inherits(cf, "try-error") || is.na(cf)) return(NA_real_)
    as.numeric(cf)
  }
  bobj <- boot::boot(data = mfz, statistic = stat_ix, R = BOOT_R)

  Boot_CI_95 <- tryCatch({
    ci <- boot::boot.ci(bobj, type = "bca", index = 1)$bca[4:5]
    fmt_ci(ci[1], ci[2])
  }, error = function(e) tryCatch({
    # Fall back to percentile CI if BCa fails (e.g. degenerate bootstrap distribution)
    ci <- boot::boot.ci(bobj, type = "perc", index = 1)$percent[4:5]
    fmt_ci(ci[1], ci[2])
  }, error = function(e2) NA_character_))

  tibble::tibble(
    Tract          = y,
    Moderator      = unname(mod_map[w]),
    R2             = fmt_r2(summary(fit)$r.squared),
    beta_std       = fmt3(est),
    partial_R2_HC3 = fmt_pr2(pr2),
    Robust_p       = apa_p(p_hc3),
    Robust_CI_95   = fmt_ci(ci_hc3[1], ci_hc3[2]),
    Boot_CI_95     = Boot_CI_95
  )
}

safe_run_one <- purrr::safely(run_one, otherwise = NULL, quiet = TRUE)
```

# Statistical Analysis

> *Methods:* Two supplementary analyses examined the robustness of the H1b moderation results. **(1) HC3 + bootstrap (Table S7):** Moderation models (PROCESS Model 1; Hayes, 2022) were re-estimated using heteroskedasticity-consistent (HC3) robust standard errors and BCa bootstrap 95% confidence intervals (5,000 resamples). An effect was considered robust when *both* criteria were simultaneously met. All models were estimated on standardised data (outcome and CTQ z-scored; moderators mean-centred). **(2) Three-way interaction (Table S8):** To directly address diagnostic comorbidity, a three-way CTQ × PTSD × mTBI interaction (PROCESS Model 3) was estimated for each tract. PTSD and mTBI status were coded as binary variables (0/1) derived from the four-group dummy variables. Tracts are those that showed significant CTQ–WM associations in H1a (FDR *p* < .05).

- **Outcome (Y):** WM tract metric (MD or AD)
- **Predictor (X):** CTQ total score
- **Moderator(s) (W):** Diagnosis group dummy (PTSD-only, mTBI-only, Comorbid; reference: HC) for Table S7; binary PTSD presence × binary mTBI presence for Table S8
- **Covariates:** Age, gender

---

# Analysis

## Table S7: HC3 Robust Standard Errors and BCa Bootstrap CIs

```{r run_s7, echo=FALSE}
set.seed(123)

# Run all tract × moderator combinations
grid <- tidyr::crossing(Tract = available_tracts, Moderator_var = available_mods)

res_list <- purrr::pmap(grid, function(Tract, Moderator_var) {
  safe_run_one(Tract, Moderator_var, data = dat)
})

# Report any models that failed
errs <- purrr::map_chr(res_list, \(x) if (is.null(x$error)) "" else x$error$message)
if (any(nzchar(errs)))
  message("Some models were skipped:\n",
          paste(unique(head(errs[nzchar(errs)], 5)), collapse = "\n"))

# Combine results and join pretty tract labels
tab <- res_list |>
  purrr::map("result") |>
  purrr::compact() |>
  dplyr::bind_rows()

tab2 <- tab %>%
  rename(Tract_raw = Tract) %>%
  mutate(Moderator = factor(Moderator, levels = c("mTBI", "PTSD", "PTSD+mTBI"))) %>%
  left_join(tract_map, by = "Tract_raw") %>%
  filter(!is.na(Group)) %>%
  mutate(Group = factor(Group, levels = c("MD", "AD"))) %>%
  arrange(Group, Tract_order, Moderator)
```

---

## Table S8: Three-way Interaction (PROCESS Model 3, Sensitivity Analysis)

```{r run_s8, echo=FALSE}
set.seed(123)

# Build binary PTSD and mTBI status flags from the existing dummy columns.
df_proc <- dat %>%
  mutate(
    ptsd_any = as.integer(`hc.ptsd.mtbi.groups.4groups_PTSD-only` == 1 |
                          `hc.ptsd.mtbi.groups.4groups_Comorbid mTBI+PTSD` == 1),
    mtbi_any = as.integer(`hc.ptsd.mtbi.groups.4groups_mTBI-only` == 1 |
                          `hc.ptsd.mtbi.groups.4groups_Comorbid mTBI+PTSD` == 1),
    ctq_c    = as.numeric(scale(ctq_totalscore)),   # z-scored (not just centred)
    ptsd_c   = ptsd_any - mean(ptsd_any, na.rm = TRUE), # mean-centred
    mtbi_c   = mtbi_any - mean(mtbi_any, na.rm = TRUE)  # mean-centred
  )

tracts_avail_proc <- intersect(tracts, names(df_proc))
if (!length(tracts_avail_proc))
  stop("run_s8: no requested tracts found in df_proc — check column names.")

term3   <- "ctq_c:ptsd_c:mtbi_c"
rows_m3 <- vector("list", length(tracts_avail_proc))

# S8 uses the same APA-style helpers as S7 (no leading zeros).
fmt_ci_s8 <- fmt_ci
fmt_b_s8  <- fmt3

for (i in seq_along(tracts_avail_proc)) {
  y   <- tracts_avail_proc[i]
  fml <- as.formula(paste0(bt(y), " ~ ctq_c * ptsd_c * mtbi_c + age + gender"))

  mf       <- model.frame(fml, data = df_proc)
  mf[[y]]  <- as.numeric(scale(mf[[y]]))   # z-score outcome per tract
  fit <- lm(fml, data = mf)

  if (!term3 %in% names(coef(fit)))
    stop(glue("run_s8: three-way term '{term3}' not found for tract '{y}' — check variable names."))

  # HC3 robust inference for the three-way term
  V     <- sandwich::vcovHC(fit, type = "HC3")
  b3    <- unname(coef(fit)[term3])
  se3   <- sqrt(diag(V))[term3]
  t3    <- b3 / se3
  p_hc3 <- 2 * pt(-abs(t3), df = df.residual(fit))
  ciH   <- lmtest::coefci(fit, vcov. = V)[term3, ]

  # BCa bootstrap CI for the three-way term
  stat3 <- function(data, idx) {
    d  <- data[idx, , drop = FALSE]
    cf <- try(coef(lm(fml, data = d))[term3], silent = TRUE)
    if (inherits(cf, "try-error") || is.null(cf) || length(cf) != 1L) return(NA_real_)
    as.numeric(cf)
  }
  bobj <- boot::boot(data = mf, statistic = stat3, R = BOOT_R)
  ciB  <- tryCatch({
    bca <- boot::boot.ci(bobj, type = "bca", index = 1)
    c(bca$bca[4], bca$bca[5])
  }, error = function(e) {
    perc <- boot::boot.ci(bobj, type = "perc", index = 1)
    c(perc$percent[4], perc$percent[5])
  })

  rows_m3[[i]] <- tibble::tibble(
    Tract          = y,
    R2             = fmt_r2(summary(fit)$r.squared),
    beta           = fmt_b_s8(b3),
    `p (HC3)`      = apa_p(p_hc3),
    `95% CI (HC3)` = fmt_ci_s8(ciH[1], ciH[2]),
    `95% CI (BCa)` = fmt_ci_s8(ciB[1], ciB[2])
  )
}

Table_S4_M3 <- dplyr::bind_rows(rows_m3) %>%
  rename(Tract_raw = Tract) %>%
  left_join(tract_map, by = "Tract_raw") %>%
  filter(!is.na(Group)) %>%
  mutate(Group = factor(Group, levels = c("MD", "AD"))) %>%
  arrange(Group, Tract_order) %>%
  transmute(
    Group, Tract = Tract_pretty, R2, beta, `p (HC3)`, `95% CI (HC3)`, `95% CI (BCa)`
  )
```

---

# Results

## Table S7

```{r display_s7, echo=FALSE}
# Build display table: beta gets leading zero restored; p loses "= " prefix (APA S7 style)
disp_s7 <- tab2 %>%
  transmute(
    Tract          = Tract_pretty,
    Moderator,
    beta           = sub("^\\.", "0.", sub("^-\\.", "-0.", beta_std)),
    `p (HC3)`      = sub("^= ", "", Robust_p),
    `95% CI (HC3)` = Robust_CI_95,
    `95% CI (BCa)` = Boot_CI_95
  )

tbl_s7 <- disp_s7 |>
  gt() |>
  tab_header(
    title = html(paste0(
      "<b>Table S7.</b> Moderation Effects with HC3 Standard Errors and Bootstrap Confidence Intervals."
    ))
  ) |>
  tab_style(
    style     = cell_text(align = "left"),
    locations = cells_title(groups = "title")
  ) |>
  cols_label(
    beta           = html("\u03b2"),
    `p (HC3)`      = html("<i>p</i> (HC3)"),
    `95% CI (HC3)` = "95% CI (HC3)",
    `95% CI (BCa)` = "95% CI (BCa)"
  ) |>
  cols_align(align = "left",   columns = Tract) |>
  cols_align(align = "center", columns = -Tract) |>
  tab_options(
    table.font.size                   = "small",
    source_notes.font.size            = "small",
    data_row.padding                  = px(2),
    heading.border.bottom.style       = "hidden",
    table.border.top.style            = "solid",
    table.border.bottom.style         = "solid",
    column_labels.border.top.style    = "solid",
    column_labels.border.bottom.style = "solid",
    table_body.border.bottom.style    = "solid"
  ) |>
  tab_source_note(source_note = html(paste0(
    "<i>Note.</i> MD, mean diffusivity; AD, axial diffusivity; AF, arcuate fasciculus; ",
    "CB, cingulum bundle; CC, corpus callosum; ILF, inferior longitudinal fasciculus; ",
    "IoFF, inferior occipito-frontal fasciculus; SLF, superior longitudinal fasciculus; ",
    "UF, uncinate fasciculus; PTSD, post-traumatic stress disorder; mTBI, mild traumatic ",
    "brain injury; HC3, heteroskedasticity-consistent; CI, confidence interval; BCa, ",
    "bias-corrected accelerated. Models estimated on standardised data (outcomes and CTQ ",
    "z-scored; moderators mean-centred). Reported: standardised interaction \u03b2, HC3 ",
    "p-values and 95% CIs, BCa bootstrap 95% CIs (5,000 resamples)."
  )))

tbl_s7

gtsave(tbl_s7,
       here("output", "tables", "supplementary", "TableS7_Moderation_HC3.html"))
```

## Table S8

```{r display_s8, echo=FALSE}
tbl_s8 <- Table_S4_M3 |>
  gt(groupname_col = "Group") |>
  tab_header(
    title = html(paste0(
      "<b>Table S8.</b> Sensitivity analysis of three-way CM × PTSD × mTBI interactions on white matter
microstructure."
    ))
  ) |>
  tab_style(
    style     = cell_text(align = "left"),
    locations = cells_title(groups = "title")
  ) |>
  tab_style(
    style     = cell_text(weight = "bold"),
    locations = cells_row_groups()
  ) |>
  cols_label(
    R2             = html("R\u00b2"),
    beta           = html("\u03b2"),
    `p (HC3)`      = html("<i>p</i> (HC3)"),
    `95% CI (HC3)` = "95% CI (HC3)",
    `95% CI (BCa)` = "95% CI (BCa)"
  ) |>
  cols_align(align = "left",   columns = Tract) |>
  cols_align(align = "center", columns = -Tract) |>
  tab_options(
    table.font.size                   = "small",
    source_notes.font.size            = "small",
    data_row.padding                  = px(2),
    heading.border.bottom.style       = "hidden",
    table.border.top.style            = "solid",
    table.border.bottom.style         = "solid",
    column_labels.border.top.style    = "solid",
    column_labels.border.bottom.style = "solid",
    table_body.border.bottom.style    = "solid",
    row_group.border.top.style        = "hidden",
    row_group.border.bottom.style     = "hidden"
  ) |>
  tab_source_note(source_note = html(paste0(
    "<i>Note.</i> MD, mean diffusivity; AD, axial diffusivity; AF, arcuate fasciculus; ",
    "CB, cingulum bundle; CC, corpus callosum; ILF, inferior longitudinal fasciculus; ",
    "IoFF, inferior occipito-frontal fasciculus; SLF, superior longitudinal fasciculus; ",
    "UF, uncinate fasciculus; PTSD, post-traumatic stress disorder; mTBI, mild traumatic ",
    "brain injury; HC3, heteroskedasticity-consistent; CI, confidence interval; BCa, ",
    "bias-corrected accelerated. Linear models tested whether the CTQ\u2013WM association ",
    "varied jointly by PTSD and mTBI status (CTQ \u00d7 PTSD \u00d7 mTBI), adjusting for ",
    "age and sex. Entries are the three-way interaction coefficient (\u03b2), model R\u00b2, ",
    "HC3 p-values with 95% CIs, and BCa bootstrap 95% CIs (5,000 resamples)."
  )))

tbl_s8

gtsave(tbl_s8,
       here("output", "tables", "supplementary", "TableS8_Moderation_3way.html"))
```
