---
title: "INTRuST CM — H2b: White Matter Mediating CTQ → Post-Concussive Symptoms"
author: "Luisa Berger"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

suppressPackageStartupMessages({
  library(here)
  library(tidyverse)
  library(gt)
})

set.seed(654321)

dat <- readr::read_csv(
  here("data", "analysis_simulated.csv"),
  show_col_types = FALSE,
  name_repair = "minimal"
)

dat$gender <- as.numeric(dat$gender)
dat$race   <- as.numeric(dat$race)

source(here("code", "utils", "process.R"), echo = FALSE)
```

```{r define_mediators, include=FALSE}
# Single source of truth for all mediators in this script.
# Names  = human-readable tract labels used in the results table.
# Values = the _1000 column names passed to run_mediation() and PROCESS.
#
# Scaling (below), analysis chunks, and the results table all derive from
# these vectors — adding or removing a mediator requires editing one line here.

md_mediators <- c(
  "CC 2"         = "t_cc2.tensor1.meandiffusivity.mean_1000",
  "CC 6"         = "t_cc6.tensor1.meandiffusivity.mean_1000"
)

ad_mediators <- c(
  "AF left"      = "t_af_left.tensor1.maxeigenvalue.mean_1000",
  "AF right"     = "t_af_right.tensor1.maxeigenvalue.mean_1000",
  "CB right"     = "t_cb_right.tensor1.maxeigenvalue.mean_1000",
  "CC 4"         = "t_cc4.tensor1.maxeigenvalue.mean_1000",
  "CC 5"         = "t_cc5.tensor1.maxeigenvalue.mean_1000",
  "IoFF right"   = "t_ioff_right.tensor1.maxeigenvalue.mean_1000",
  "SLF I left"   = "t_slf.i_left.tensor1.maxeigenvalue.mean_1000",
  "SLF I right"  = "t_slf.i_right.tensor1.maxeigenvalue.mean_1000",
  "SLF II left"  = "t_slf.ii_left.tensor1.maxeigenvalue.mean_1000",
  "SLF II right" = "t_slf.ii_right.tensor1.maxeigenvalue.mean_1000"
)
```

```{r scale_wm_columns, include=FALSE}
# WM diffusion values rescaled × 1,000,000 for numerical stability in PROCESS (suffix "_1000").

all_mediators <- c(md_mediators, ad_mediators)
base_cols     <- sub("_1000$", "", all_mediators)

missing_cols <- setdiff(base_cols, names(dat))
if (length(missing_cols) > 0) {
  stop("scale_wm_columns: the following expected columns are absent from the data:\n  ",
       paste(missing_cols, collapse = "\n  "))
}

for (i in seq_along(all_mediators)) {
  dat[[all_mediators[i]]] <- dat[[base_cols[i]]] * 1000000
}
```

```{r helpers, include=FALSE}
# parse_section(): extract numeric values from a named section of PROCESS output.
parse_section <- function(out, keyword, min_vals = 4) {
  # Anchor search after the "DIRECT AND INDIRECT EFFECTS" header to avoid false matches.
  effects_hdr <- which(grepl("direct",   out, ignore.case = TRUE) &
                       grepl("indirect", out, ignore.case = TRUE))
  search_from <- if (length(effects_hdr) > 0) tail(effects_hdr, 1L) else 1L

  local_idx <- grep(keyword, out[search_from:length(out)], fixed = TRUE)
  if (length(local_idx) == 0) {
    # Fallback: search full output in case section header format differs
    local_idx <- grep(keyword, out, fixed = TRUE)
    if (length(local_idx) == 0)
      stop("parse_section: '", keyword, "' not found in PROCESS output.")
    search_from <- 1L
  }
  abs_idx <- local_idx[1L] + search_from - 1L

  for (i in seq(abs_idx + 1L, min(abs_idx + 20L, length(out)))) {
    line  <- out[i]
    if (is.na(line) || nchar(trimws(line)) == 0) next
    nums  <- suppressWarnings(as.numeric(strsplit(trimws(line), "\\s+")[[1]]))
    valid <- !is.na(nums)
    if (sum(valid) >= min_vals) return(nums[valid])
  }
  context <- out[seq(abs_idx, min(abs_idx + 21L, length(out)))]
  stop("parse_section: no data row found in '", keyword, "'.\nLines scanned:\n",
       paste(context, collapse = "\n"))
}

# run_mediation(): run PROCESS Model 4 once, print output inline, return key effects as tibble.
# Running each model once avoids doubling bootstrap computation (~10,000 iterations × 12 models).
run_mediation <- function(dat, mediator, save = 0) {

  # Run PROCESS once and capture all console output as a character vector.
  out <- capture.output(
    process(
      data    = dat,
      y       = "rpq13_total",
      x       = "ctq_totalscore",
      m       = mediator,
      cov     = c("age", "gender"),
      effsize = 1,
      total   = 1,
      stand   = 1,
      boot    = 10000,
      modelbt = 1,
      seed    = 654321,
      save    = save,
      model   = 4
    )
  )

  cat(out, sep = "\n")  # re-print inline so the Analysis section output is unchanged

  # Parse the three effect sections and return as a tibble (invisible so it
  # doesn't print a second time in the chunk output).
  total  <- parse_section(out, "Total effect of X on Y")
  direct <- parse_section(out, "Direct effect of X on Y")
  # Indirect effects: "Indirect effect(s) of X on Y" appears multiple times in
  # PROCESS output, and so does "BootSE". Some of those pairings correspond to
  # the M-model bootstrap table (constant ~1,250 when mediator is ×1e6-scaled)
  # rather than the actual indirect-effects bootstrap CI table.
  # Robust strategy: scan valid section pairings from LAST to FIRST; accept the
  # first one whose data row has all 4 values with |x| < 10 (indirect effects
  # and their bootstrap CIs are always small; M-model constants are not).
  indir  <- local({
    indir_lines <- grep("Indirect effect", out, ignore.case = TRUE)
    indir_lines <- indir_lines[!grepl("Completely standardized", out[indir_lines], ignore.case = TRUE)]
    if (length(indir_lines) == 0)
      stop("run_mediation: 'Indirect effect' section not found in PROCESS output.")
    for (sec_idx in rev(indir_lines)) {
      window  <- out[seq(sec_idx + 1L, min(sec_idx + 5L, length(out)))]
      if (!any(grepl("BootSE", window, fixed = TRUE))) next
      hdr_rel <- grep("BootSE", window, fixed = TRUE)[1L]
      start   <- sec_idx + hdr_rel + 1L
      for (i in seq(start, min(start + 10L, length(out)))) {
        line  <- out[i]
        if (is.na(line) || nchar(trimws(line)) == 0) next
        nums  <- suppressWarnings(as.numeric(strsplit(trimws(line), "\\s+")[[1]]))
        valid <- !is.na(nums)
        vals  <- nums[valid]
        if (length(vals) >= 4L) {
          if (all(abs(vals) < 10)) return(vals)
          break
        }
        break
      }
    }
    stop("run_mediation: no plausible indirect-effects row found in PROCESS output.")
  })

  invisible(tibble(
    b_c      = total[7],  t_c      = total[3],  p_c      = total[4],
    b_cprime = direct[7], t_cprime = direct[3], p_cprime = direct[4],
    b_ab     = indir[1],  se_ab    = indir[2],  llci     = indir[3],
    ulci     = indir[4]
  ))
}
```

# Statistical Analysis

> *Methods:* Mediation analyses explored whether WM microstructure mediates the associations between CM severity and PTSD or post-concussive symptom severity. All mediation analyses were conducted using R PROCESS macro (version 4, model 4; Hayes, 2022), which employs non-parametric bootstrapping with *N* = 10,000 permutations to generate 95% bias-corrected confidence intervals. The dMRI measures that showed significant associations with CM severity in H1a (FDR *p* < .05) were included as mediators. CM severity (CTQ total score) was the independent variable in all models, and RPQ-13 score was the dependent variable.

- **Outcome (Y):** RPQ-13 total score (post-concussive symptoms)
- **Predictor (X):** CTQ total score
- **Mediator (M):** WM tract metric (MD or AD)
- **Covariates:** Age, gender

---

# Analysis

## Mean Diffusivity (MD)

### MD — CC 2

```{r mediation_MD_CC2, echo=FALSE}
# save = 1 retains bootstrap samples for this primary MD model.
r_md_cc2 <- run_mediation(dat, md_mediators[["CC 2"]], save = 1)
```

### MD — CC 6

```{r mediation_MD_CC6, echo=FALSE}
r_md_cc6 <- run_mediation(dat, md_mediators[["CC 6"]])
```

---

## Axial Diffusivity (AD)

### AD — AF left

```{r mediation_AD_AF_left, echo=FALSE}
r_ad_af_left <- run_mediation(dat, ad_mediators[["AF left"]])
```

### AD — AF right

```{r mediation_AD_AF_right, echo=FALSE}
r_ad_af_right <- run_mediation(dat, ad_mediators[["AF right"]])
```

### AD — CB right

```{r mediation_AD_CB_right, echo=FALSE}
r_ad_cb_right <- run_mediation(dat, ad_mediators[["CB right"]])
```

### AD — CC 4

```{r mediation_AD_CC4, echo=FALSE}
r_ad_cc4 <- run_mediation(dat, ad_mediators[["CC 4"]])
```

### AD — CC 5

```{r mediation_AD_CC5, echo=FALSE}
r_ad_cc5 <- run_mediation(dat, ad_mediators[["CC 5"]])
```

### AD — IoFF right

```{r mediation_AD_IoFF_right, echo=FALSE}
r_ad_ioff_right <- run_mediation(dat, ad_mediators[["IoFF right"]])
```

### AD — SLF I left

```{r mediation_AD_SLF_I_left, echo=FALSE}
r_ad_slf_i_left <- run_mediation(dat, ad_mediators[["SLF I left"]])
```

### AD — SLF I right

```{r mediation_AD_SLF_I_right, echo=FALSE}
r_ad_slf_i_right <- run_mediation(dat, ad_mediators[["SLF I right"]])
```

### AD — SLF II left

```{r mediation_AD_SLF_II_left, echo=FALSE}
r_ad_slf_ii_left <- run_mediation(dat, ad_mediators[["SLF II left"]])
```

### AD — SLF II right

```{r mediation_AD_SLF_II_right, echo=FALSE}
r_ad_slf_ii_right <- run_mediation(dat, ad_mediators[["SLF II right"]])
```

---

# Results

```{r results_table, echo=FALSE}
# N is the number of participants with complete RPQ data (PROCESS uses listwise deletion)
n_obs <- sum(!is.na(dat$rpq13_total))

# Assemble all stored mediation results; each r_* object is a one-row tibble
# returned (invisibly) by run_mediation() above — no re-computation needed.
# Tract labels come from the mediator vector names (single source of truth).
med_results <- bind_rows(
  r_md_cc2          %>% mutate(metric = "MD", tract = names(md_mediators["CC 2"])),
  r_md_cc6          %>% mutate(metric = "MD", tract = names(md_mediators["CC 6"])),
  r_ad_af_left      %>% mutate(metric = "AD", tract = names(ad_mediators["AF left"])),
  r_ad_af_right     %>% mutate(metric = "AD", tract = names(ad_mediators["AF right"])),
  r_ad_cb_right     %>% mutate(metric = "AD", tract = names(ad_mediators["CB right"])),
  r_ad_cc4          %>% mutate(metric = "AD", tract = names(ad_mediators["CC 4"])),
  r_ad_cc5          %>% mutate(metric = "AD", tract = names(ad_mediators["CC 5"])),
  r_ad_ioff_right   %>% mutate(metric = "AD", tract = names(ad_mediators["IoFF right"])),
  r_ad_slf_i_left   %>% mutate(metric = "AD", tract = names(ad_mediators["SLF I left"])),
  r_ad_slf_i_right  %>% mutate(metric = "AD", tract = names(ad_mediators["SLF I right"])),
  r_ad_slf_ii_left  %>% mutate(metric = "AD", tract = names(ad_mediators["SLF II left"])),
  r_ad_slf_ii_right %>% mutate(metric = "AD", tract = names(ad_mediators["SLF II right"]))
) %>%
  # Show metric label (MD/AD) only on the first row of each group, matching
  # the manuscript table layout where the label functions as a row-span.
  mutate(metric_label = if_else(!duplicated(metric), metric, ""))

tbl_med <- med_results %>%
  select(metric_label, tract,
         b_c, t_c, p_c,
         b_cprime, t_cprime, p_cprime,
         b_ab, se_ab, llci, ulci) %>%
  gt() %>%
  tab_header(
    title = html(paste0(
      "<b>Table S4.</b> Mediation results of childhood maltreatment severity ",
      "on post-concussive symptom severity (<i>N</i> = ", n_obs, ")."
    ))
  ) %>%
  tab_style(style = cell_text(align = "left"), locations = cells_title(groups = "title")) %>%
  # ── column spanners (outer → inner for indirect effect) ─────────────────────
  tab_spanner(
    label   = html("<b>Total effect</b> (<i>c</i>)"),
    columns = c(b_c, t_c, p_c)
  ) %>%
  tab_spanner(
    label   = html("<b>Direct effect</b> (<i>c</i>&#x2032;)"),
    columns = c(b_cprime, t_cprime, p_cprime)
  ) %>%
  # Inner spanner: CI sub-header nested inside the indirect effect spanner
  tab_spanner(
    label   = "Percentile bootstrap\n95% confidence interval",
    columns = c(llci, ulci),
    id      = "ci_span"
  ) %>%
  # Outer indirect effect spanner that wraps b_ab, se_ab, and the CI spanner
  tab_spanner(
    label    = html("<b>Indirect effect</b> (<i>ab</i>)"),
    columns  = c(b_ab, se_ab),
    spanners = "ci_span"
  ) %>%
  # ── column labels ────────────────────────────────────────────────────────────
  cols_label(
    metric_label = "",
    tract        = "",
    b_c          = html("<i>b</i><sub>c</sub>"),
    t_c          = html("<i>t</i>-value"),
    p_c          = html("<i>p</i>-value"),
    b_cprime     = html("<i>b</i><sub>c&#x2032;</sub>"),
    t_cprime     = html("<i>t</i>-value"),
    p_cprime     = html("<i>p</i>-value"),
    b_ab         = html("<i>b</i><sub>ab</sub>"),
    se_ab        = "SE",
    llci         = "Lower",
    ulci         = "Upper"
  ) %>%
  # ── number formatting ────────────────────────────────────────────────────────
  fmt_number(columns = -c(metric_label, tract), decimals = 3) %>%
  # Strip leading zeros (0.xxx → .xxx; −0.xxx → −.xxx) and flag p = 0.000
  # as "< .001" per APA convention (p cannot equal zero).
  text_transform(
    locations = cells_body(columns = c(p_c, p_cprime)),
    fn = function(x) {
      dplyr::case_when(
        x == "0.000" ~ "< .001",
        TRUE         ~ sub("^0\\.", ".", x)
      )
    }
  ) %>%
  text_transform(
    locations = cells_body(columns = -c(metric_label, tract, p_c, p_cprime)),
    fn = function(x) {
      x <- sub("^0\\.",       ".",       x)   # 0.xxx  → .xxx
      x <- sub("^\u22120\\.", "\u2212.", x)   # −0.xxx → −.xxx
      x
    }
  ) %>%
  # ── conditional bold ─────────────────────────────────────────────────────────
  # Metric labels and tract names
  tab_style(
    style     = cell_text(weight = "bold"),
    locations = cells_body(columns = c(metric_label, tract))
  ) %>%
  # Significant p-values (total and direct effects)
  tab_style(
    style     = cell_text(weight = "bold"),
    locations = cells_body(columns = c(p_c, p_cprime), rows = p_c < 0.05)
  ) %>%
  # Indirect effect row when bootstrap CI excludes zero (significant mediation)
  tab_style(
    style     = cell_text(weight = "bold"),
    locations = cells_body(columns = c(b_ab, se_ab, llci, ulci), rows = llci > 0)
  ) %>%
  # ── alignment ────────────────────────────────────────────────────────────────
  cols_align(align = "left",   columns = c(metric_label, tract)) %>%
  cols_align(align = "center", columns = -c(metric_label, tract)) %>%
  # ── APA-style borders ────────────────────────────────────────────────────────
  tab_options(
    table.font.size                   = "small",
    source_notes.font.size            = "small",
    data_row.padding                  = px(2),
    heading.border.bottom.style       = "hidden",
    table.border.top.style            = "solid",
    table.border.bottom.style         = "solid",
    column_labels.border.top.style    = "solid",
    column_labels.border.bottom.style = "solid",
    table_body.border.bottom.style    = "solid"
  ) %>%
  # ── table note ───────────────────────────────────────────────────────────────
  tab_source_note(
    source_note = html(paste0(
      "<i>Note.</i> MD, mean diffusivity; AD, axial diffusivity; CC, corpus callosum; ",
      "AF, arcuate fasciculus; CB, cingulum bundle; IoFF, inferior occipito-frontal fasciculus; ",
      "SLF, superior longitudinal fasciculus. ",
      "The association between childhood maltreatment severity and post-concussive symptom ",
      "severity was partially mediated by mean diffusivity and axial diffusivity in ",
      "various white matter tracts. All analyses were adjusted for age and gender as covariates."
    ))
  )

tbl_med

gtsave(tbl_med,
       here("output", "tables", "supplementary", "TableS4_Mediation_RPQ.html"))

write.csv(
  med_results %>% select(-metric_label),
  here("output", "results", "main", "Mediation_WM_RPQ_results.csv"),
  row.names = FALSE
)
```
